---
title: "Danish Kings, W.10"
author: "Caroline-Marie Jacobsen"
date: "07-03-2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The task here is to load your Danish Monarchs csv into R using the `tidyverse` toolkit, calculate and explore the kings' duration of reign with pipes `%>%` in `dplyr` and plot it over time.

## Load the kings

Make sure to first create an `.Rproj` workspace with a `data/` folder where you place either your own dataset or the provided `kings.csv` dataset.

1.  Look at the dataset that are you loading and check what its columns are separated by? (hint: open it in plain text editor to see)

List what is the

separator: ","

2.  Create a `kings` object in R with the different functions below and inspect the different outputs.

-   `read.csv()`
-   `read_csv()`
-   `read.csv2()`
-   `read_csv2()`

```{r eval = TRUE}
# FILL IN THE CODE BELOW and review the outputs

library(tidyverse)

kings1 <- read.csv("data/Kongerækken.csv")

kings2 <- read_csv("data/Kongerækken.csv")

kings3 <- read.csv2("data/Kongerækken.csv")

kings4 <- read_csv2("data/Kongerækken.csv")
```

Answer:  

1. Which of these functions is a `tidyverse` function? Read data with it below into a `kings` object  
- Kings 2 and 4 are tidyverse functions, while kings 1 and 3 are base R functions.

2. What is the result of running `class()` on the `kings` object created with a tidyverse function.  
- When running class(kings) it shows that the object kings is a tibble (tbl) data frame.

3.  How many columns does the object have when created with these different functions?  
- To find the number of columns I type ncol(kings). It will be 1 or 8 columns. In which case the right result is the one with 8. 

4.  Show the dataset so that we can see how R interprets each column  
- To show the dataset I type glimpse(kings). It will show me how R interprects each column. The results can be seen below. 

```{r eval = TRUE}
# COMPLETE THE BLANKS BELOW WITH YOUR CODE, then turn the 'eval' flag in this chunk to TRUE.

kings <- read_csv("data/Kongerækken.csv")

class(kings)

ncol(kings)

glimpse(kings)
```

## Calculate the duration of reign for all the kings in your table

You can calculate the duration of reign in years with `mutate` function by subtracting the equivalents of your `startReign` from `endReign` columns and writing the result to a new column called `duration`. But first you need to check a few things:

-   Is your data messy? Fix it before re-importing to R
-   Do your start and end of reign columns contain NAs? Choose the right strategy to deal with them: `na.omit()`, `na.rm=TRUE`, `!is.na()`

Create a new column called `duration` in the kings dataset, utilizing the `mutate()` function from tidyverse. Check with your group to brainstorm the options.

```{r}
#The column "Antal_regeringsår" did already exist in my dataset. Here I am just showing how I would create the column and then just overriding the already existing column named "Antal_regeringsår". 

kings <- kings %>%
  mutate(Antal_regeringsår = Regering_Slut - Regering_Start)
 

```

## Calculate the average duration of reign for all rulers

Do you remember how to calculate an average on a vector object? If not, review the last two lessons and remember that a column is basically a vector. So you need to subset your `kings` dataset to the `duration` column. If you subset it as a vector you can calculate average on it with `mean()` base-R function. If you subset it as a tibble, you can calculate average on it with `summarize()` tidyverse function. Try both ways!

-   You first need to know how to select the relevant `duration` column. What are your options?
-   Is your selected `duration` column a tibble or a vector? The `mean()` function can only be run on a vector. The `summarize()` function works on a tibble.
-   Are you getting an error that there are characters in your column? Coerce your data to numbers with `as.numeric()`.
-   Remember to handle NAs: `mean(X, na.rm=TRUE)`

```{r}
#I am finding the mean() by using base R function. By writing "kings$Antal_regeringsår", I choose only to find the average in this specific column. 

mean(kings$Antal_regeringsår, na.rm = TRUE)
```
```{r}
#and with summarise

kings %>% 

summarise(mean(Antal_regeringsår, na.rm = TRUE))

```

## How many and which kings enjoyed a longer-than-average duration of reign?

You have calculated the average duration above. Use it now to `filter()` the `duration` column in `kings` dataset. Display the result and also count the resulting rows with `count()`

```{r}
#I am finding the kings who had a longer than average duration of reign. 

kings_longer_than_average <- kings %>%
  filter(Antal_regeringsår > mean(Antal_regeringsår, na.rm =TRUE))

glimpse(kings_longer_than_average)

```
```{r}
#and count 

count(kings_longer_than_average)
```


## How many days did the three longest-ruling monarchs rule?

-   Sort kings by reign `duration` in the descending order. Select the three longest-ruling monarchs with the `slice()` function
-   Use `mutate()` to create `Days` column where you calculate the total number of days they ruled
-   BONUS: consider the transition year (with 366 days) in your calculation!

```{r}
#I will use the slice_max function to find the 3 longest-ruling monarchs. Setting the number = 3 will give me the result. 

kings %>% 
  slice_max(Antal_regeringsår, n = 3)

```
```{r }
#I am creating the new column "Antal_dage"

kings <- kings %>%
  mutate(Antal_dage = Antal_regeringsår * 365)

#And running it to show the result 

kings %>% 
  slice_max(Antal_dage, n = 3)

```

## Challenge: Plot the kings' duration of reign through time

What is the long-term trend in the duration of reign among Danish monarchs? How does it relate to the historical violence trends ?

-   Try to plot the duration of reign column in `ggplot` with `geom_point()` and `geom_smooth()`
-   In order to peg the duration (which is between 1-99) somewhere to the x axis with individual centuries, I recommend creating a new column `midyear` by adding to `startYear` the product of `endYear` minus the `startYear` divided by two (`startYear + (endYear-startYear)/2`).
-   Now you can plot the kings dataset, plotting `midyear` along the x axis and `duration` along y axis
-   BONUS: add a title, nice axis labels to the plot and make the theme B&W and font bigger to make it nice and legible!

```{r}
# YOUR CODE

```

```{r}


```

And to submit this rmarkdown, knit it into html. But first, clean up the code chunks, adjust the date, rename the author and change the `eval=FALSE` flag to `eval=TRUE` so your script actually generates an output. Well done!
